---
title: "Homework2"
author: "Emily Surrell"
format: html
editor: visual
---

# Task 1: Basic Vector Practice

Suppose we have data from a medical experiment on blood pressure.

1.  Create two vectors. One vector corresponding to the pre-treatment values for subjects 1 through 20 and one vector corresponding to the post-treatment values for subjects 1 through 20.

```{r}

#assign values to pre_treatment
pre_treatment <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

#assign values to post_treatment
post_treatment <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

#view results
pre_treatment
post_treatment
```

2.  Assign names to the vector elements using the paste() function.

```{r}

#assign names
names(pre_treatment) <- paste("Subject", 1:20, sep = "_")
names(post_treatment) <- paste("Subject", 1:20, sep = "_")

#view results
pre_treatment
post_treatment

```

3. Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements.

```{r}

#calculate change in blood pressure
blood_pressure_change <- pre_treatment - post_treatment

#view results
blood_pressure_change

```

4. Calculate the average decrease in blood pressure accross all patients.

```{r}
mean(blood_pressure_change)
```

5. Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

```{r}

which(blood_pressure_change > 0)

```

6. Subset the vector of differences to only return those that have a positive change.

```{r}

#subset the vector
blood_pressure_decreased <- blood_pressure_change[blood_pressure_change > 0]

#view results
blood_pressure_decreased

```

7. Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}

mean(blood_pressure_decreased)

```

