---
title: "Homework 2"
author: "Emily Surrell"
format: html
editor: visual
---

# Task 1: Basic Vector Practice

Suppose we have data from a medical experiment on blood pressure.

1. Create two vectors. One vector corresponding to the pre-treatment values for subjects 1 through 20 and one vector corresponding to the post-treatment values for subjects 1 through 20.

```{r}

#assign values to pre_treatment
pre_treatment <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

#assign values to post_treatment
post_treatment <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

#view results
pre_treatment
post_treatment
```

2. Assign names to the vector elements using the paste() function.

```{r}

#assign names
names(pre_treatment) <- paste("Subject", 1:20, sep = "_")
names(post_treatment) <- paste("Subject", 1:20, sep = "_")

#view results
pre_treatment
post_treatment

```

3. Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements.

```{r}

#calculate change in blood pressure
blood_pressure_change <- pre_treatment - post_treatment

#view results
blood_pressure_change

```

4. Calculate the average decrease in blood pressure accross all patients.

```{r}
mean(blood_pressure_change)
```

5. Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

```{r}

which(blood_pressure_change > 0)

```

6. Subset the vector of differences to only return those that have a positive change.

```{r}

#subset the vector
blood_pressure_decreased <- blood_pressure_change[blood_pressure_change > 0]

#view results
blood_pressure_decreased

```

7. Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}

mean(blood_pressure_decreased)

```

# Task 2: Basic Data Frame Practice

Continue the previous example.

1. Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp.

```{r}

#create data frame
blood_pressure <- data.frame(patient = 1:20, pre_bp = pre_treatment, post_bp = post_treatment, diff_bp = blood_pressure_change)

#view results
blood_pressure

```

2. Return only rows where the diff_bp column is negative.

```{r}

blood_pressure[blood_pressure$diff_bp < 0,]

```

3. Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120.

```{r}

blood_pressure$lessthan120 <- (blood_pressure$post_bp < 120)

```

4. Finally, print the data frame out nicely.

```{r}

knitr::kable(blood_pressure)

```

# Task 3: List Practice

Continue the previous example. Suppose we now have another experiment where the "treatment" was actually a placebo.

1. Create a new data frame for the placebo data.

```{r}

#create initial data frame to store data
placebo <- data.frame(patient = 1:10, pre_bp = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130), post_bp = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124))

#calculate blood pressure changes
placebo$diff_bp <- placebo$pre_bp - placebo$post_bp

#flag when post-treatment blood pressure is less than 120
placebo$lessthan120 <- (placebo$post_bp < 120)

#view results
placebo

```

2. Create and store a list with the two data frames as elements.

```{r}

#create list
experiment <- list(treatment = blood_pressure, placebo = placebo)

#view results
experiment

```

3. Access the first list element using three different types of syntax.

```{r}

experiment[1]
experiment[[1]]
experiment$treatment

```

4. In one line, access the placebo data frame, pre_bp column.

```{r}

experiment$placebo$pre_bp

```

# Task 4: Control Flow Practice

Continue the previous example. Suppose we want to charactize the post-treatment or post-placebo blood pressure measurement as optimal (<=120), borderline (120 < bp <= 130), and high (> 130).

1. First, create a new column in each data frame called status.

```{r}

#create status columns
experiment$treatment$status <- character(20)
experiment$placebo$status <- character(10)

#view results
experiment

```

2. For the non-placebo data frame, create the status column's values.

```{r}

#fill in values for status
for (patient in experiment$treatment$patient){
  if (experiment$treatment$post_bp[patient] <= 120){
    experiment$treatment$status[patient] <- "optimal"
  } else if (experiment$treatment$post_bp[patient] <= 130){
    experiment$treatment$status[patient] <- "borderline"
  } else {
    experiment$treatment$status[patient] <- "high"
  }
}

#view results

experiment$treatment
```

3. Repeat for the placebo data frame.

```{r}

#fill in values for status
for (patient in experiment$placebo$patient){
  if (experiment$placebo$post_bp[patient] <= 120){
    experiment$placebo$status[patient] <- "optimal"
  } else if (experiment$placebo$post_bp[patient] <= 130){
    experiment$placebo$status[patient] <- "borderline"
  } else {
    experiment$placebo$status[patient] <- "high"
  }
}

#view results

experiment$placebo

```

