---
title: "Homework2"
author: "Caroline Hutchings"
date: today
format: html
editor: visual
---

# Vectors

## Create Vectors

Create two vectors for blood pressure measurements. One vector is for pre-treatment measurements and the second vector is for post-treatment measurements.

```{r}
pre_treat <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_treat <- c( 114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

## Name Vector Elements

Add names to each of the blood pressure measurements for both the pre- treatment blood pressure measurements and the post-treatment blood pressure measureents

```{r}
names(pre_treat) <- paste("Subject", 1:20, sep = "_")
pre_treat

names(post_treat) <- paste("Subject", 1:20, sep = "_")
post_treat
```

## Vector Manipulations

Determine if the bp treatment is effective by finding the difference in bp using before treatment minus after treatment.

```{r}
change_bp <- pre_treat-post_treat
change_bp
str(change_bp)
```

### Mean Change

Determine the average change in bp after treatment.

```{r}
mean(change_bp)
```

### Identify Elements

Identify each subject with a bp that decreased after treatment, or a positive change in bp since we have pre-treatment bp - post treatment bp

```{r}
?which()
which(change_bp>0)
```

### Subset Vector

Create a vector that only includes those subjects with a decrease in bp after treatment.

```{r}
positive <- change_bp[change_bp>0]
length(positive)


```

#### Average Decrease

Determine the average decrease in bp.

```{r}
mean(positive)
```

# Data Frame

Create a data frame with patient, pre treatment bp, post treatment bp, and the differnce in bp

```{r}
patient_info <- data.frame(patient=names(pre_treat), pre_bp=pre_treat, post_bp = post_treat, diff_bp=change_bp, row.names=NULL)
patient_info
```

### Subset the Data Frame

Select rows from the data frame where the difference in bp is negative. In other words select the rows where the pre-treatment bp is lower than the post-treatment bp.

```{r}
neg_bp <- patient_info[patient_info$diff_bp<0,]
neg_bp
```

### Adding a New Column

Add a column that compares the post-treatment bp to 120 and returns true when the bp is less than 120.

```{r}
patient_info$true <- patient_info$post_bp<120
patient_info

```

### Print the Data Frame

```{r}
knitr::kable(patient_info)
```
