---
title: "HW2_SM"
format: html
editor: visual
---

## Task 1: Basic Vector Practice Create Pre- and Post-procedure Blood Pressure Measurement Vectors.

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Assign names to each vector's elements.

```{r}
subject_names <- paste("Subject", 1:20, sep = "_")
```

Calculate change in blood pressure following treatment

```{r}
diff_bp <- (pre_bp - post_bp)
```

Average Decrease in BP across all patients

```{r}
mean(diff_bp)
```

Identify patients that experienced a drop in blood pressure following treatment

```{r}
which(diff_bp > 0)
```
Subset patients with blood pressure drop into a vector
```{r}
bp_drop <- diff_bp[diff_bp > 0]
```

Calculate average blood pressure drop following treatment
```{r}
mean(bp_drop)
```

## Task 2: Basic Data Frame Practice 
## Task 3: List Practice 
## Task 4: Control Flow Practice
