---
title: "Homework 2"
author: Cameron Mullaney
date: 8/29/2026
format: html
editor: visual
---

### 1. Basic Vector Practice

Here, I create the named numeric vectors given in the assignment

```{r}
## Measurements
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

##Subject IDs
names(pre) <- paste("Subject", 1:20, sep = "_")
names(post) <- paste("Subject", 1:20, sep = "_")
```

Here, I create some new vectors based on data we have

```{r}
bp_change <- pre-post
cat("Average blood pressure change:", mean(bp_change))
names(which(bp_change > 0))
bp_decrease <- which(bp_change > 0)
cat("Average blood pressure decrease (among those who decreased):", mean(bp_decrease))
```

### 2. Basic Data Frame Practice

Here I'm compiling these vectors into one data frame. `rownames(bp_df) <- NULL` resets the indices to numbers, since we have patient ID in its own column

```{r}
bp_df <- data.frame(
  patient = names(pre),
  pre_bp = pre,
  post_bp = post,
  diff_bp = bp_change)

rownames(bp_df) <- NULL
```

Taking a look at `diff_bp` to see which patients' blood pressure increased:

```{r}
subset(bp_df, bp_df$diff_bp < 0)
```

Here, I am creating a new column `low_bp` to categorize `post_bp`

```{r}
bp_df$low_bp <- ifelse(bp_df$post_bp <120, TRUE, FALSE)
```

Now, let's take a look at the final table:

```{r}
knitr::kable(bp_df)
```

### 3. List Practice

Combining the treatment and placebo data frames into one list

```{r}
placebo <- data.frame(
  patient = paste("Subject", 1:10, sep = "_"),
  pre_bp = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130),
  post_bp = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
)
placebo$diff_bp <- placebo$pre_bp - placebo$post_bp
placebo$low_bp = ifelse(placebo$post_bp <120, TRUE, FALSE)

all <- list(treatment = bp_df, placebo = placebo)
```

Three ways to access treatment data:

```{r}
all[1]
all[["treatment"]]
all$treatment
```

Accessing the placebo group `pre_bp`

```{r}
all$placebo["pre_bp"]
```

### 4. Control Flow Practice

Two nested for loops:

`i` iterates through the two data frames, and `j` iterates through the rows of each data frame.

For each row, a "status" is set based on post-treatment blood pressure.

```{r}
all$treatment["status"] <- "optimal"
all$placebo["status"] <- "optimal"

## Iterate through treatment & placebo

for (i in 1:2){
  
  ## Iterate through rows of each
  
  for (j in 1:nrow(all[[i]])){
    if ((all[[i]][["post_bp"]][j]) > 130) {
      all[[i]][["status"]][j] = "high"
      } else if(all[[i]][["post_bp"]][j] > 120) {
       all[[i]][["status"]][j] = "borderline"
       }
  }
}
```

Taking a look at the two tables:

```{r}
cat("Treatment\n"); knitr::kable(all["treatment"])
cat("Placebo\n"); knitr::kable(all["placebo"])
```

All looks good! Blood pressures were properly categorized
