---
title: "Homework 2"
author: "ST 558 Group 5"
date: today
format: html
editor: visual
---

# 1. Basic Vector practice

### Create and manipulate

1.  Create two vectors. One vector corresponding to the pre-treatment values for subjects 1 through 20 and one vector corresponding to the post-treatment values for subjects 1 through 20.

```{r}
pre_treat <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 
               140, 137, 131, 120, 128, 139, 135)

post_treat <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 
                122, 106, 106, 124, 102, 117, 113)
```

2.  Add subject IDs to each of the blood pressure measurements for both the pre-treatment blood pressure measurements and the post-treatment blood pressure measurements

```{r}
names(pre_treat) <- paste("Subject", 1:20, sep = "_")
pre_treat

names(post_treat) <- paste("Subject", 1:20, sep = "_")
post_treat
```

3.  Determine if the treatment is effective by finding the difference in blood pressure before treatment and after treatment.

```{r}
change_bp <- pre_treat - post_treat
change_bp

```

### Calculate

4.  Determine the average change in bp after treatment.

```{r}
cat("Average blood pressure change:", mean(change_bp))
```

### Use `which()`

5.  Identify each subject with a bp that decreased after treatment, or a positive change in bp since we have pre-treatment bp - post treatment bp

```{r}
which(change_bp > 0)
```

### Subset

6.  Create a vector that only includes those subjects with a decrease in bp after treatment.

```{r}
positive <- change_bp[change_bp > 0]
positive


```

7.  Determine the average decrease in bp.

```{r}
mean(positive)
```

# 2. Basic Data Frame practice

1.  Create a data frame with columns for patient ID, pre treatment bp, post treatment bp, and the difference in bp

```{r}
patient_info <- data.frame(patient = names(pre_treat), 
                           pre_bp = pre_treat, 
                           post_bp = post_treat, 
                           diff_bp = change_bp, 
                           row.names = NULL)

patient_info
```

### Subset

2.  Select rows from the data frame where the difference in bp is negative. In other words select the rows where the pre-treatment bp is lower than the post-treatment bp.

```{r}
neg_bp <- patient_info[patient_info$diff_bp < 0,]
neg_bp
```

### Add a New Column

3.  Add a column that compares the post-treatment bp to 120 and returns true when the bp is less than 120.

```{r}
patient_info$true <- patient_info$post_bp < 120
patient_info

```

### Print

4.  Print the data frame

```{r}
knitr::kable(patient_info)
```

# 3. List practice

### New Data Frame

1.  Create a new data frame with only 10 subjects that contains all of the same variables as the original data frame.

```{r}
## Pre- & post- "treatment" vectors for placebo group
pre_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

#Combine into data frame, with same calculated columns 
patient_placebo <- data.frame(patient_names = paste("Subject", 1:10, sep = "_"),
                              pre_bp = pre_placebo, 
                              post_bp = post_placebo, 
                              diff_bp = pre_placebo-post_placebo, 
                              true = post_placebo < 120)

## Result: 
patient_placebo
```

### List creation

2.  Create a list that contains both of data frames, treatment and placebo.

```{r}
patient_list <- list(treatment = patient_info, placebo = patient_placebo)
patient_list
```

### List access

3.  Access the first element of the list in 3 different ways

```{r}
patient_list[1]
patient_list[[1]]
patient_list$treatment
```

4.  Access the placebo data frame from the list and pre_bp column.

```{r}
patient_list$placebo$pre_bp
```

# 4. Control flow practice

1.  Create a new column in each data frame called status and identify the status rating for each post treatment bp and store in the data frame in the list.

```{r}
#create the status column for treatment data frame
patient_list[[1]]$status <- character(20)

#create status column for placebo data frame
patient_list[[2]]$status <- character(10)

patient_list
```

2.  Fill the treatment group's status column with ratings from subjects' post treatment blood pressure.

```{r}
for (i in 1:20) {
  if (patient_list[[1]][i, "post_bp"] <= 120) {
    patient_list[[1]][i,"status"] <- "Optimal"
  } 
  else if (patient_list[[1]][i, "post_bp"] <= 130) {
    patient_list[[1]][i,"status"] <- "Borderline"
  } 
  else {
    patient_list[[1]][i,"status"] <- "High"
  }
}
```

3.  Repeat for placebo group.

```{r}
for (i in 1:10) {
  if (patient_list[[2]][i, "post_bp"] <= 120) {
    patient_list[[2]][i,"status"] <- "Optimal"
  } 
  else if (patient_list[[2]][i, "post_bp"] <= 130) {
    patient_list[[2]][i,"status"] <- "Borderline"
  } 
  else {
    patient_list[[2]][i,"status"] <- "High"
  }
}
```

View each table with the new status column

```{r}
cat("Treatment\n"); knitr::kable(patient_list["treatment"])
cat("Placebo\n"); knitr::kable(patient_list["placebo"])
```
