---
title: "HW4_CM"
format: html
---

```{r}
library(tidyverse)
library(readxl)
library(RSQLite)
```

## Task 1: Conceptual Questions

1.  Nested function (baseR)

```{r}
iris |>
  as.tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length > 1.55) |>
  arrange(Species)
```

2.  "Long format" refers to the more common data layout, where each row corresponds to a single observation or data point, often (but not always) separated by time. "Wide format" data features multiple grouped observations in a single row, united by a common variable (city, week, team, etc.). Data in this format is typically more difficult to analyze, and usually requires transformation into long format data, which is preferred for statistical analysis.

## Task 2: Data Handling

#### Glass Data

```{r}
glassColNames <- c('id', 'ri', 'na', 'mg', 'al', 'si', 'k', 'ca', 'ba', 'fe', 'type_of_glass')
glass <- read_csv('https://www4.stat.ncsu.edu/~online/datasets/glass.data', 
                  col_names = glassColNames)
```

3.  Type of glass
4.  Filtering

```{r}
## 3 ##
glass |>
  mutate(type_of_glass = factor(type_of_glass, 1:7, 
                                c('building_windows_float_processed',
                                  'building_windows_non_float_processed',
                                  'vehicle_windows_float_processed', 
                                  'vehicle_windows_non_float_processed',
                                  'containers',
                                  'tableware',
                                  'headlamps'))) |>
## 4 ##
  filter(fe < 0.2, type_of_glass %in% c('tableware', 'headlamps'))
```

#### Yeast Data

```{r}
yeastColNames<-c('seq', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class')

yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = yeastColNames, show_col_types = FALSE)
yeast
```

5.  Drop columns
6.  Add group means

```{r}
yeast |>
## 5 ##
  select(-c(seq, nuc)) |>
## 6 ##
  group_by(class) |>
  mutate(across(where(is.numeric), mean, .names = "mean_{.col}"))

```

#### Wine Data

```{r}
## Read in each wine sheet

white <- read_excel("../HW_3/white-wine.xlsx", sheet = 1, .name_repair = "universal") |>
  mutate(color = "white")
  
red <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
                  delim = ";", name_repair = "universal") |>
  mutate(color = "red")

## Bind together
wine <- dplyr::bind_rows(white, red)
```

7.  Filter to high-quality wines
8.  Sort by quality (descending)
9.  Select certain columns
10. Calculated alcohol mean & standard deviation grouped by quality

```{r}
wine |>
## 7 ##
  filter(quality > 6.5, alcohol < 13.2)|>
## 8 ##
  arrange(desc(quality))|>
## 9 ## 
  select(contains("acid"), alcohol, color, quality) |>
## 10 ##
  group_by(quality) |>
  mutate("alc.mean" = mean(alcohol)) |>
  mutate("alc.sd" = sd(alcohol))

```
