---
title: "HW4_SM"
author: Sofia Mariano
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  Rewrite the following nested function call using baseR's chaining operator:

```{r}
# original code:
library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

# Re-written code:
library(tidyverse)
iris_tbl<- as_tibble(iris) |>
  dplyr::select(starts_with("Petal"), Species) |>
  dplyr::filter(Petal.Length < 1.55) |>
  dplyr::arrange(Species)
iris_tbl
```

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

- Long format data: each row is one observation and each column is one variable
- Wide format data: more than one observation may be present in a given row.
- Long format data is generally preferred for performing statistical analysis.

# Task 2:
## Glass Data
Prepare session to read-in data sets
```{r}
library(tidyverse)
```

```{r}
# Read in original code from HW3
glass.data <- readr::read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = c("ID","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Glass")
                       )
glass.data

# Overwrite Type_of_Glass variable with string labels
glass.data |>
  
```

## Yeast Data

## Wine Data

# Task 3: Database Practice
