---
title: "HW4_SM"
author: Sofia Mariano
format: html
editor: visual
---
Prepare session to read-in data sets and perform various functions
```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(readxl)
library(DBI)
library(RSQLite)
```

# Task 1: Conceptual Questions

1.  Rewrite the following nested function call using baseR's chaining operator:

```{r}
# original code:
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

# Re-written code:
iris_tbl<- as_tibble(iris) |>
  dplyr::select(starts_with("Petal"), Species) |>
  dplyr::filter(Petal.Length < 1.55) |>
  dplyr::arrange(Species)
iris_tbl
```

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

- Long format data: each row is one observation and each column is one variable
- Wide format data: more than one observation may be present in a given row.
- Long format data is generally preferred for performing statistical analysis.

# Task 2:
## Glass Data
Read in original code from HW3
```{r}
glass.data <- readr::read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = c("ID","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type_of_Glass")
                       )
glass.data
```

Modify Glass Dataset code
```{r}
glass.data |>
  dplyr::mutate(Type_of_Glass = factor(Type_of_Glass, 1:7, c("building_windows_float_processed", "build
ing_windows_non_float_processed", "vehicle_windows_float_processed", "vehicle_windows_non_float_processed", "containers", "tableware", "headlamps"))
) |>
  dplyr::filter(Fe < 0.2 & Type_of_Glass %in% c("tableware", "headlamp"))
glass.data

```

## Yeast Data
Read-in original code from HW3
```{r}
yeast.data <- readr::read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                       fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                                (c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")
                                  ))
)
yeast.data
```

Perform data modifications
```{r}
yeast.data |>
  dplyr::select(mcg:vac, class) |>
  dplyr::group_by(class) |>
  dplyr::mutate(across(where(is.numeric), mean, .names = "{.col}_mean"))
```

## Wine Data
Read-in original code from HW3
```{r}
#Read in and modify white wine data set
white_wine.data <- readxl::read_excel("../HW_3/white-wine.xlsx",
                              sheet = 1
                              )
wine_var_names <- readxl::read_excel("../HW_3/white-wine.xlsx",
                              sheet = 2
                           )
colnames(white_wine.data) <- t(wine_var_names) # reading and correcting column names
white_wine.data$type <- "white"# adding wine type

# Read in and modify read wine data set
red_wine.data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                            delim = ";",show_col_types = FALSE)
colnames(red_wine.data) <- t(wine_var_names)#Fix Column Names
red_wine.data$type <-"red"#Add type column

# Combine red and white wine into single data set
wine.data <- dplyr::bind_rows(white_wine.data,red_wine.data)
wine.data
```

Modify data set
```{r}
wine.data |>
  dplyr::filter(quality > 6.5 & alcohol < 13.2) |>
  dplyr::arrange(desc(quality)) |>
  dplyr::select(1:3,alcohol, type, quality) |>
  dplyr::group_by(quality) |>
  dplyr::mutate(across(alcohol, list(mean = mean, sd = sd)))
```

# Task 3: Database Practice
Read in Lahman data base
```{r}
# Connect database
lahman <- DBI::dbConnect(RSQLite::SQLite(), "lahman.db")

# look at tables within database
DBI::dbListTables(lahman)
```

Return data from Teams table in 2015
```{r}
# Return data using dplyr
dplyr::tbl(lahman, "Teams") |>
  dplyr::filter(yearID == 2015) |>
  collect()

# Return data using SQL
dplyr::tbl(lahman, sql("select *
                        From Teams
                       where yearID = 2015")) |>
  collect()
```

Create hall of fame player table
```{r}
# Look at original hall of fame table
dplyr::tbl(lahman, "HallOfFame")

# Create hall of fame table
dplyr::tbl(lahman, "HallOfFame") |>
  dplyr::filter(inducted == 'Y') |>
  dplyr::select(playerID, yearID, category) |>
  collect()
```

Combine with People information
```{r}
# Look at original People table
dplyr::tbl(lahman, "People")

# Join with hall of fame table
hall_of_fame <- dplyr::left_join(
  dplyr::tbl(lahman, "HallOfFame") |>
    dplyr::filter(inducted == 'Y') |>
    dplyr::select(playerID, yearID, category)
  ,
  dplyr::tbl(lahman, "People") |>
    dplyr::select(playerID, nameFirst, nameLast)
  ,
  by = c("playerID")
  ) |>
collect()
hall_of_fame
```

Create manager table
```{r}
# Look at original Managers table
dplyr::tbl(lahman, "Managers")

# Create managers table
manager <- dplyr::tbl(lahman, "Managers") |>
  dplyr::select(playerID, G, W, L) |>
  dplyr::group_by(playerID) |>
  dplyr::summarize(G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
    ) |>
  collect() |>
  dplyr::mutate(Career_W_Percentage = Total_W/G_managed) |>
  dplyr::arrange(desc(Total_W/G_managed))
manager
```

Join manager and hall of fame tables
```{r}
dplyr::inner_join(manager, hall_of_fame, by = c("playerID"))
```

Disconnect from database
```{r}
dbDisconnect(lahman)
```
