---
title: "HW4_Group"
author: "ST 558 Group 5"
date: 9/15/2026
format: html
---
Prepare session to perform tasks
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(readr)
library(DBI)
library(RSQLite)
```

# Task 1: Conceptual Questions
1. Rewrite the following nested call function using **baseR**'s chaining operator:

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

```{r}
as_tibble(iris) |>
  dplyr::select(starts_with("Petal"), Species) |>
  dplyr::filter(Petal.Length < 1.55) |>
  dplyr::arrange(Species)
```

2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

* Long format data refers to structured data where each row is one observation and each column is one variable.

* Wide format data refers to structured data with more than one observation present in a single row.  These observations are often united by a common variable.

* For statistical analysis, long format data is generally preferred.

# Task 2:
## Glass Data

