---
title: "Homework 4"
author: "Emily Surrell"
date: 9/14/2026
format: html
---

Bring in libraries needed for this assignment.

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```

### Task 1: Conceptual Questions

1.  Rewrite the following nested function call using **baseR**’s chaining operator:

```         
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
```

Using chaining operator:

```{r}
as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

    Long format data is structured such that each observation is a row and each variable is a column. Wide format data has more than one observation in a row.

    Although wide format data can make for a nicer display, long format data is generally preferred for statistical analysis.

### Task 2: Chaining Practice

#### Glass Data

Read in glass data.

```{r}
glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type_of_glass'))
```

Clean glass data with chaining.

```{r}
glass <- glass |>
  mutate(Type_of_glass = factor(Type_of_glass, 1:7, c('building_windows_float_processed','building_windows_non_float_processed','vehicle_windows_float_processed','vehicle_windows_non_float_processed','containers','tableware','headlamps'))) |>
  filter(Fe < 0.2, (Type_of_glass %in% c('tableware', 'headlamps')))

#print results
glass
```

#### Yeast Data

Read in yeast data.

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))
```

Clean yeast data with chaining.

```{r}
yeast <- yeast |>
  select(!c(seq_name,nuc)) |>
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean=mean, median=median), .names = "{.col}_{.fn}"))

#print results
yeast
```

#### Wine Data

Read in and combine wine data.

```{r}
#white wine
white <- read_excel("../HW_3/white-wine.xlsx", sheet = 1)
colnamesdata <- read_excel("../HW_3/white-wine.xlsx", sheet = 2)
colnames(white) <- t(colnamesdata)
white$type <- "white"

#red wine
red <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
colnames(red) <- t(colnamesdata)
red$type <- "red"

#combine
wine <- dplyr::bind_rows(white, red)
```

Clean wine data with chaining.

```{r}
wine <- wine |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol),
        sd_alcohol = sd(alcohol))

#print results
wine

```

### Task 3: Database Practice
