---
title: "Homework 4"
author: Cameron Mullaney
date: today
format: html
warning: false
results: hold
---

First, loading in the necessary packages:

```{r}
library(tidyverse)
library(readxl)
library(RSQLite)
```

## Task 1: Conceptual Questions

1.  Nested function in baseR

```{r}
iris |>
  as.tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length > 1.55) |>
  arrange(Species)
```

2.  "Long format" refers to the more common data layout, where each row corresponds to a single observation or data point, often (but not always) separated by time. "Wide format" data features multiple grouped observations in a single row, united by a common variable (city, week, team, etc.). Data in this format is typically more difficult to analyze, and usually requires transformation into long format data, which is preferred for statistical analysis.

## Task 2: Data Handling

#### Glass Data

Pulling in, naming, and manipulating a csv from the web

```{r}
glassColNames <- c('id', 'ri', 'na', 'mg', 'al', 'si', 'k', 'ca', 'ba', 'fe', 'type_of_glass')
glass <- read_csv('https://www4.stat.ncsu.edu/~online/datasets/glass.data', 
                  col_names = glassColNames)
glass
```

3.  Providing names to `type_of_glass` column, and treating it as a factor
4.  Filtering by `fe` and `type_of_glass`

```{r}
## 3 ##
glass |>
  mutate(type_of_glass = factor(type_of_glass, 1:7, 
                                c('building_windows_float_processed',
                                  'building_windows_non_float_processed',
                                  'vehicle_windows_float_processed', 
                                  'vehicle_windows_non_float_processed',
                                  'containers',
                                  'tableware',
                                  'headlamps'))) |>
## 4 ##
  filter(fe < 0.2, type_of_glass %in% c('tableware', 'headlamps'))
```

#### Yeast Data

Loading in a data table from the web and providing column names

```{r}
yeastColNames<-c('seq', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class')

yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = yeastColNames, show_col_types = FALSE)
yeast
```

5.  Drop columns `seq` and `nuc`
6.  Add columns for the means of each numeric variable, grouped by `class`

```{r}
yeast |>
## 5 ##
  select(-c(seq, nuc)) |>
## 6 ##
  group_by(class) |>
  mutate(across(where(is.numeric), 
                list(mean = mean, median = median),
                .names = "{.col}_{.fn}"))

```

#### Wine Data

Loading in data from a .xlsx file (white wine) and from a *semicolon-separated values* file (red wine)

Then including a column to specify wine color, and stacking the two data sets using `dplyr::bind_rows`

```{r}
## Read in each wine sheet

white <- read_excel("../HW_3/white-wine.xlsx", sheet = 1, .name_repair = "universal") |>
  mutate(color = "white")
  
red <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
                  delim = ";", name_repair = "universal") |>
  mutate(color = "red")

## Bind together
wine <- dplyr::bind_rows(white, red)
```

7.  Filter to high-quality wines (using `quality`)
8.  Sort by `quality` (descending)
9.  Select acid, `color`, and `quality` columns
10. Calculate `alcohol` mean & standard deviation grouped by `quality`

```{r}
wine |>
## 7 ##
  filter(quality > 6.5, alcohol < 13.2)|>
## 8 ##
  arrange(desc(quality))|>
## 9 ## 
  select(contains("acid"), alcohol, color, quality) |>
## 10 ##
  group_by(quality) |>
  mutate("alc.mean" = mean(alcohol)) |>
  mutate("alc.sd" = sd(alcohol))

```

## Task 3: Database Practice

11. Connect to database in repo, and view list of tables

```{r}
con <- dbConnect(RSQLite::SQLite(), "lahman.db")
dbListTables(con)
```

12. Using `tbl()` and `dplyr` to pull 2015 teams

```{r}
tbl(con, "Teams")|>
  filter(yearID == 2015)
```

13. Using `tbl()` and `sql()` to pull 2015 teams

```{r}
tbl(con, sql(
  "SELECT *
  FROM `Teams`
  WHERE (`yearID` = 2015)")
  )
```

14. All players in the Hall of Fame

```{r}
HOF <- tbl(con, "HallOfFame")|>
  filter(inducted == "Y")|>
  select(playerID, yearID, category)

HOF
```

15. Add in their names from the 'People' table

```{r}
HOF <- HOF|>
  left_join((tbl(con, "People") |> select(playerID, nameFirst, nameLast)), 
             by = "playerID")|>
  collect()
HOF
```

16. Manager Stats

```{r}
Mgr <- tbl(con, "Managers")|>
  select(playerID, G, W, L) |>
  group_by(playerID)|>
  summarize(totG = sum(G, na.rm = TRUE),
            totW = sum(W, na.rm = TRUE),
            totL = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(Wpct = (totW / totG)*100)|>
  arrange(desc(Wpct))

Mgr
```

17. Combine 15 & 16 to list all HOF members who managed at least one game.

```{r}
HOF |>
  inner_join(Mgr, by = "playerID")
```

Finally, disconnect from the database.

```{r}
dbDisconnect(con)
```
