---
title: "Homework 4"
author: "Caroline Hutchings"
date: today
date-format: "MM/DD/YYYY"
format: pdf
editor: visual
---

## Conceptual Questions

library(tidyverse) arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length \< 1.55), Species)

1.  Rewrite using baseR chaining operator

```{r}
library(tidyverse)
iris |> 
  as_tibble() |> 
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length<1.55) |> 
  arrange(Species)
```

2.  What is meant by long format data and wide format data?
    1.  Long format is a data table organization where each row is one observation and each column is a variable. Repeated measures are stored in additional rows. This is the ideal format when working with structured data for statistical analysis.
    2.  Wide format is where there can be multiple observations per row. Repeated measures are stored in additional columns instead of additional rows. Sometimes this makes the data easier to read in a table format.
3.  Which do we generally prefer for statistical analysis? We generally prefer the long format to do analyses on. This enables us to do analysis on the variables and include each observation separately, not necessarily each individual though as some rows may be repeated measures on the same individual.

## Task 2

### Glass Data

Read in the glass data to a tibble. Change the variable Type to be a factor variable with character names and filter the rows to include only the tableware and headlamps type and Fe less than 0.2.

```{r}
#Read in the glass data to a tibble
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names=c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type"))
glass_data

#Change the variable Type to be a factor and filter the rows

glass_data <- glass_data |> 
  mutate(Type = factor(Type, levels=c(1,2,3,4,5,6,7), labels= c('building_windows_float_processed', 'building_windows_non_float_processed', 'vehicle_windows_float_processed', 'vehicle_windows_non_float_processed', 'containers', 'tableware', 'headlamps'))) |>
  filter(Fe<0.2, Type %in% c('tableware' ,'headlamps'))
glass_data
```

### Yeast Data

Read Yeast data into a tibble and add column names. Remove the columns 'Sequence' and 'nuc' by chaining. Add a mean and median column for each numerical variable. Group each mean and median for each variable by the class variable.

```{r}
# Read data into tibble

yeast_data <- read_fwf('https://www4.stat.ncsu.edu/online/datasets/yeast.data', fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3), col_names=c("Sequence", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class") ))
yeast_data

#Remove sequence and nuc columns and add mean and median for each column grouped by class

yeast_data <- yeast_data |> 
  select(mcg:vac, class) |> 
  group_by(class) |> 
  mutate(across(where(is.numeric), c(mean=mean, median=median), .names='{.fn}_{.col}')) 

yeast_data
```

### Wine Data

Load the wine data and combine the red and white wine datasets using the streamlined version.

```{r}
library(readxl)
library(here)
wine_names <- read_excel(here("HW_3", "white-wine.xlsx"), sheet = 2)

white_wine_data <- read_excel(here("HW_3", "white-wine.xlsx")) |>
  setNames(wine_names$Variables) |>
  mutate(type = "White")

red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",delim = ";") |>
  setNames(wine_names$Variables) |>
  mutate(type = "Red")

wine_data <- bind_rows(white_wine_data, red_wine)
wine_data
```

Filter the data to look only at higher quality wines and reasonable alcohol content,

```{r}
wine_data <- wine_data |> 
  filter(quality>6.5, alcohol<13.2) |> 
  arrange(desc(quality)) |> 
  select(contains('acid'), alcohol, type, quality) |> 
  group_by(quality) |> 
  mutate(across(alcohol, c(mean=mean, std= sd), .names='{.fn}_{.col}'))

```
