---
title: "HW4_Group"
author: "ST 558 Group 5"
date: 9/15/2026
format: html
---
Prepare session to perform tasks
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(readr)
library(DBI)
library(RSQLite)
```

# Task 1: Conceptual Questions
1. Rewrite the following nested call function using **baseR**'s chaining operator:

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

```{r}
as_tibble(iris) |>
  dplyr::select(starts_with("Petal"), Species) |>
  dplyr::filter(Petal.Length < 1.55) |>
  dplyr::arrange(Species)
```

2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

* Long format data refers to structured data where each row is one observation and each column is one variable.

* Wide format data refers to structured data with more than one observation present in a single row.  These observations are often united by a common variable.

* For statistical analysis, long format data is generally preferred.

# Task 2:
## Glass Data
Read in original Hw3 code
```{r}
glass <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type_of_Glass'))
```

Modify data set to change **glass type** variable output and keep observations where **Fe** < 0.2 and **glass type** is "tableware" or "headlamp"

```{r}
glass <- glass|>
  dplyr::mutate(Type_of_Glass = factor(Type_of_Glass, 1:7, c('building_windows_float_processed','building_windows_non_float_processed','vehicle_windows_float_processed','vehicle_windows_non_float_processed','containers','tableware','headlamps'))) |>
  dplyr::filter(Fe < 0.2, (Type_of_Glass %in% c('tableware', 'headlamps'))
         )
# Print results
glass
```

## Yeast Data
Read in original HW3 code
```{r}
yeast <- readr::read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
  col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))

```

Modify dataset to remove **seq_name** and **nuc** columns and ad columns corresponding to the mean and media of each numeric variable at each **class** grouping.
```{r}
yeast <- yeast |> 
  dplyr::select(!c(seq_name,nuc)) |> 
  dplyr::group_by(class) |> 
  dplyr::mutate(across(where(is.numeric), c(mean=mean, median=median), .names='{.fn}_{.col}')) 

# Print results
yeast
```


## Wine Data
Read in original HW3 code
```{r}
#white wine
white <- readxl::read_excel("../HW_3/white-wine.xlsx", sheet = 1)
colnamesdata <- readxl::read_excel("../HW_3/white-wine.xlsx", sheet = 2)
colnames(white) <- t(colnamesdata)
white$type <- "white"

#red wine
red <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
colnames(red) <- t(colnamesdata)
red$type <- "red"

#combine
wine <- dplyr::bind_rows(white, red)
```

Modify data set to look at **quality** > 6.5, **alcohol** < 13.2, sort data from highest to lowest **quality**, print variables that contain only **acid**, **alcohol**, **type**, and **quality**, and determine the mean and standard deviation of **alcohol** based on each setting of **quality**
```{r}
wine <- wine |> 
  dplyr::filter(quality > 6.5, alcohol < 13.2) |> 
  dplyr::arrange(desc(quality)) |> 
  dplyr::select(contains('acid'), alcohol, type, quality) |> 
  dplyr::group_by(quality) |> 
  dplyr::mutate(across(alcohol, c(mean=mean, std= sd), .names='{.fn}_{.col}'))

# Print results
wine
```

# Task 3: Database Practice
Connect to database and view tables
```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), "lahman.db")
dbListTables(con)
```

Return Teams table from 2015 using dplyr
```{r}
dplyr::tbl(con, "Teams")|>
  dplyr::filter(yearID == 2015)|>
  collect()
```

Returm Teams table from 2015 using sql
```{r}
dplyr::tbl(con, sql(
  "SELECT *
  FROM `Teams`
  WHERE (`yearID` = 2015)")
  )|>
  collect()
```

From **HallOfFame** table, return players inducted, year voted in, and category
```{r}
dplyr::tbl(con, "HallOfFame") |>
  dplyr::filter(inducted == 'Y') |>
  dplyr::select(playerID, yearID, category) |>
  collect()
```

Combine **HallOfFame** table with  **People** table to have **nameFirst** and **nameLast** returned in addition to the table above
```{r}
hall_of_fame <- dplyr::left_join(
  dplyr::tbl(con, "HallOfFame") |>
    dplyr::filter(inducted == 'Y') |>
    dplyr::select(playerID, yearID, category)
  ,
  dplyr::tbl(con, "People") |>
    dplyr::select(playerID, nameFirst, nameLast)
  ,
  by = c("playerID")
  ) |>
collect()

# Print results
hall_of_fame
```

Return **playerID**, **G**, **W**, and **L** columns from the **Managers** table using chaining
```{r}
Managers <- dplyr::tbl(con, "Managers") |>
  dplyr::select(playerID, G, W, L) |>
  dplyr::group_by(playerID) |>
  dplyr::summarize(G_managed = sum(G, na.rm = TRUE),
                   Total_W = sum(W, na.rm = TRUE),
                   Total_L = sum(L, na.rm = TRUE)) |>
  dplyr::collect() |>
  dplyr::mutate(WinPercent = Total_W/G_managed) |>
  dplyr::arrange(desc(WinPercent))

# Print results
Managers
```

Combine results from previous two questions to return information from people that managed a team and were inducted into the hall of fame
```{r}
dplyr::inner_join(Managers, hall_of_fame, by = "playerID")
```

Disconnect from the Database
```{r}
dbDisconnect(con)
```

