---
title: "Homework 5"
author: "Group 5"
date: 9/23/2026
format: html
---

Bring in libraries needed for this assignment.

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
library(psych)
```

### Task 1: Querying an API and a Tidy-Style Function

Use `GET()` from the `httr` package to return information about a topic in the news.

```{r}
getresult <- GET("https://newsapi.org/v2/everything?q=Dolly%20Parton&from=2026-09-01&language=en&pageSize=50&apiKey=2c43d850c570494389b660ffaf1757db")
```

Parse what is returned.

```{r}
parsedresult <- fromJSON(rawToChar(getresult$content))
newsdata <- as_tibble(parsedresult$articles)

#print results
newsdata
```

Write a function that allows the user to query this API.

```{r}
querynews <- function(subject, time, key){
  
  getresult <- GET(paste0("https://newsapi.org/v2/everything?q=", 
                          gsub(" ", "%20", subject), 
                          "&from=", 
                          time, 
                          "&language=en&pageSize=50&apiKey=", 
                          key))
  
  parsedresult <- fromJSON(rawToChar(getresult$content))
  return(as_tibble(parsedresult$articles))
}
```

Examples using the function:

```{r}
key <- "2c43d850c570494389b660ffaf1757db"

querynews("Nashville International Airport", "2026-09-05", key)
querynews("Miley Cyrus", "2026-09-20", key)
```

### Task 2: Summarizing Roller Coaster Data

Read in roller coaster data and convert variables to factors as needed.

```{r}
rollercoasters <- read_csv("http://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

rollercoasters <- rollercoasters |>
  mutate(across(c(AgeGroup, Scale, `Census Region`, Type, Design, `Inversions?`), as.factor))

```

Look at how the data is stored.

```{r}
str(rollercoasters)
```

Look at a summary of the data.

```{r}
psych::describe(rollercoasters)
```

Count missing values.

```{r}
colSums(is.na(rollercoasters))
```

One-way contingency table using `table()`:

```{r}
table(rollercoasters$AgeGroup, useNA="always")
```

The above table shows that 376 roller coasters are in the newest age group.

Two-way contingency table using `table()`:

```{r}
table(rollercoasters$AgeGroup,
      rollercoasters$Scale,
      useNA = "always")
```

The above table shows that 14 roller coasters are in the newest age group and are considered Kiddie roller coasters.

Three-way contingency table using `table()`:

```{r}
table(rollercoasters$AgeGroup,
      rollercoasters$Scale,
      rollercoasters$Type,
      useNA = "always")
```

The above tables show that there are 4 Kiddie roller coasters that are in the older age group and that are made of Steel.

Next, we will create a table showing counts between AgeGroup and Scale, limited to when Type is "Wood". We will do this using two different methods:

1.  Conditional two-way table using `table()`, by sub-setting the data first:

```{r}
filtered <- filter(rollercoasters, Type == "Wood")

table(filtered$AgeGroup,
      filtered$Scale,
      useNA = "always")
```

2.  Conditional two-way table using `table()`, by creating a three-way table and then sub-setting it:

```{r}
three_way <- table(rollercoasters$AgeGroup,
      rollercoasters$Scale,
      rollercoasters$Type,
      useNA = "always")

three_way[,,"Wood"]
```

Two-way contingency table using `group_by()`, `summarize()`, and `pivot_wider()`:

```{r}
rollercoasters |>
  group_by(AgeGroup, Scale) |>
  summarize(count = n()) |>
  pivot_wider(names_from = Scale, values_from = count)
```

Stacked bar graph:

```{r}
stackedbar <- ggplot(data = rollercoasters, aes(x=Scale, fill=AgeGroup))
stackedbar + geom_bar() +
   labs(x = "Scale", 
        y = "Number of Roller Coasters", 
        fill = "Age Group", 
        title = "Roller Coasters by Scale and Age Group")
```

Side-by-side bar graph:

```{r}
sidebysidebar <- ggplot(data = rollercoasters, aes(x=Scale, fill=AgeGroup))
sidebysidebar + geom_bar(position = "dodge") +
   labs(x = "Scale", 
        y = "Number of Roller Coasters", 
        fill = "Age Group", 
        title = "Roller Coasters by Scale and Age Group")
```
