---
title: "Homework 5"
author: "Caroline Hutchings"
format: html
editor: visual
---

## Task 1 Querying an API

```{r}
#install.packages("httr")
#install.packages("jsonlite")
library(httr)
library(jsonlite)
library(dplyr)

el_nino <- GET("https://newsapi.org/v2/everything?q=el%20nino&from=2026-08-20&sortBy=publishedAt&apiKey=0c0a26f9583b46fba0b935e610e59bf6")

el_nino

names(el_nino)

parsed <- fromJSON(rawToChar(el_nino$content))

names(parsed)

el_nino_news <- as_tibble(parsed$articles)

el_nino_news
```

### Create an API Function

```{r}
get_news <- function(subject, time_frame, key) {
  url_news <- paste0("https://newsapi.org/v2/everything?q=",
    subject,
    "&from=",
    time_frame,
    "&language=en",
    "&pageSize=50",
    "&apiKey=",
    key)
  #use the url to get the data
  data <- GET(url_news)
  #parse the JSON data 
  parsed <- fromJSON(rawToChar(data$content))
  #create a tibble from the parsed data
  tibble_data <- as_tibble(parsed$articles)
  #return the tibble with all the data
  return(tibble_data)
}

api_key <- "0c0a26f9583b46fba0b935e610e59bf6"

#Use function
get_news("tesla", "2026-08-20", api_key)

get_news('NCAA', "2026-08-20", api_key)

get_news("Economy", "2026-08-20", api_key)
```

## Task 2 Summarizing Data

Read in data to roller coaster tibble and remove unwanted variables.

```{r}
library(readr)
library(tidyverse)

roller_coaster <- read_csv('https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv')

names(roller_coaster)

#remove unwanted variables
roller_coaster <- roller_coaster |> 
  select(-Latitude, -Longitude, -Boundary)
```

Identify qualitative variables that should be factors - these variables should not have a large number of levels. Therefore, Park and Make are not appropriate variables to change to factors. Change remaining variables to factors.

```{r}

#identify qualitative variables that should be factors
unique_counts <- roller_coaster |> summarize(across(c(AgeGroup, Scale, Park, `Census Region`, Type, Design, `Inversions?`, Make), n_distinct))

#unique_counts

#force appropriate qualitative variables to be factors
roller_coaster <- roller_coaster |> 
  mutate(across(c(AgeGroup, Scale, `Census Region`, Type, Design, `Inversions?`), as.factor))
```

Identify missing values for each variable in the data set.

```{r}

colSums(is.na(roller_coaster))

```

Variables with missing values are:

- Drop has 128 missing values
- Top Speed has 74 missing values
- Duration has 66 missing values
- Make has 36 missing values
- Length has 33 missing values
- Max Height has 8 missing values

### Categorical Variables

Create a one-way contingency table for Type. The table shows there are 530 steel and 105 wood roller coasters.

```{r}
table(roller_coaster$Type)
```

Create a two-way contingency table for Type (Wood or Steel) and Inversions (Yes or No). Most of the roller coasters with inversions are steel, of the 170 coasters with inversions, 166 are steel and 4 are wood.

```{r}
table(roller_coaster$Type, roller_coaster$`Inversions?`)

prop.table(table(roller_coaster$Type, roller_coaster$`Inversions?`))
```

Create a three-way contingency table for Type (Steel, Wood), Inversions (Yes, No) and Age Group (Newest, Recent, Older). Only the newest wood roller coasters have any inversions. The older and recent wood roller coasters do not have any inversions. In addition, only 8 of the older steel roller coasters have any inversions. Indicating that perhaps advancing technology is required for inversions and especially inversions for wood roller coasters.

```{r}
table(roller_coaster$Type, roller_coaster$`Inversions?`, roller_coaster$AgeGroup)
```

Create a conditional two-way table using filter().

```{r}
steel <- roller_coaster |> 
  filter(Type=='Steel') 

table(steel$AgeGroup, steel$`Inversions?`)
```

Create a conditional table by sub setting a three-way contingency table.

```{r}
table(roller_coaster$Type, roller_coaster$AgeGroup, roller_coaster$`Inversions?`)["Steel",,]
```

Create a two-way contingency table using dplyr functions groupby and summarize.

```{r}
roller_coaster |> 
  group_by(Type, `Inversions?`) |> 
  drop_na(Type, `Inversions?`) |> 
  summarize(count=n()) |> 
  pivot_wider(names_from = Type, values_from = count)
```

Create a stacked and side by side bar plot

```{r}
library(ggplot2)
#stacked bar chart
ggplot(roller_coaster, aes(x=Type, fill=AgeGroup))+ geom_bar() + labs(title = "Roller Coaster Age Group by Type", x='Type', y='Count')

#side by side bar chart
ggplot(roller_coaster, aes(x=Type, fill=AgeGroup))+ geom_bar(position='dodge') + labs(title = "Roller Coaster Age Group by Type", x='Type', y='Count')
```
