---
title: "Homework 5"
author: "Caroline Hutchings"
format: html
editor: visual
---

## Task 1 Querying an API

```{r}
#install.packages("httr")
install.packages("jsonlite")
library(httr)
library(jsonlite)
library(dplyr)

el_nino <- GET("https://newsapi.org/v2/everything?q=el%20nino&from=2026-08-20&sortBy=publishedAt&apiKey=0c0a26f9583b46fba0b935e610e59bf6")

el_nino

names(el_nino)

parsed <- fromJSON(rawToChar(el_nino$content))

names(parsed)

el_nino_news <- as_tibble(parsed$articles)

el_nino_news
```

### Create an API Function

```{r}
get_news <- function(subject, time_frame, key) {
  url_news <- paste0("https://newsapi.org/v2/everything?q=",
    subject,
    "&from=",
    time_frame,
    "&language=en",
    "&pageSize=50",
    "&apiKey=",
    key)
  #use the url to get the data
  data <- GET(url_news)
  #parse the JSON data 
  parsed <- fromJSON(rawToChar(data$content))
  #create a tibble from the parsed data
  tibble_data <- as_tibble(parsed$articles)
  #return the tibble with all the data
  return(tibble_data)
}

api_key <- "0c0a26f9583b46fba0b935e610e59bf6"

#Use function
get_news("tesla", "2026-08-20", api_key)

get_news('NCAA', "2026-08-20", api_key)

get_news("Economy", "2026-08-20", api_key)
```

## Task 2 Summarizing Data

Read in data to roller coaster tibble and remove unwanted variables.

```{r}
library(readr)
library(tidyverse)

roller_coaster <- read_csv('https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv')

names(roller_coaster)

#remove unwanted variables
roller_coaster <- roller_coaster |> 
  select(-Latitude, -Longitude, -Boundary)
```

Identify qualitative variables that should be factors - these variables should not have a large number of levels. Therefore, Park and Make are not appropriate variables to change to factors. Change remaining variables to factors.

```{r}

#identify qualitative variables that should be factors
unique_counts <- roller_coaster |> summarize(across(c(AgeGroup, Scale, Park, `Census Region`, Type, Design, `Inversions?`, Make), n_distinct))

#unique_counts

#force appropriate qualitative variables to be factors
roller_coaster <- roller_coaster |> 
  mutate(across(c(AgeGroup, Scale, `Census Region`, Type, Design, `Inversions?`), as.factor))
```

Identify missing values for each variable in the data set.

```{r}

colSums(is.na(roller_coaster))

```

Variables with missing values are:

- Drop has 128 missing values
- Top Speed has 74 missing values
- Duration has 66 missing values
- Make has 36 missing values
- Length has 33 missing values
- Max Height has 8 missing values
