---
title: "Homework 5"
format: html
author: Cameron Mullaney
date: today
warning: false
---

Pulling in some relevant packages

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
```

## Task 1: Querying an API

#### 1. Using `GET()`

```{r}
turtles <- GET("https://newsapi.org/v2/everything?q=turtle&from=2026-09-15&language=en&apiKey=a86e58b83b9b4212a6fc377f18b228f9")
turtles
```

#### 2. Parsing the result

This is a very complicated object, looking through and find an article list

```{r}
names(turtles)
article.list <- fromJSON(rawToChar(turtles$content))
article.list <- as.tibble(select(article.list$articles, 2:8))
article.list
```

#### 3. Querying with a function

##### a. Function writing

```{r}
newsQuery <- function(subject, date, key){
  url <- paste0("https://newsapi.org/v2/everything?q=",
                gsub(" ", "%20", subject), 
                "&from=", date,
                "&language=en&apiKey=", key)
  query <- GET(url)
  
  art.list <- fromJSON(rawToChar(query$content))
  art.list <- as.tibble(select(art.list$articles, 2:8))
  return(art.list)
}
```

##### b. Function testing

```{r}
test1 <- newsQuery(subject = "rats", 
                  date = "2026-09-10", 
                  key = "a86e58b83b9b4212a6fc377f18b228f9")
test1
```

```{r}
test2 <- newsQuery(subject = "climate change", 
                   date = "2026-09-05", 
                   key = "a86e58b83b9b4212a6fc377f18b228f9")
test2
```

## Task 2 - Roller Coaster Data

Read in the data!

```{r}
coaster <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv", name_repair = "universal")
```

Now, cleaning the data, treating factors as such. Dropping `Latitude`, `Longitude`, and `Boundary`.

```{r}
coaster <- coaster |> 
  rename("Inversions" = "Inversions.")|>
  rename("NumInversions" = "..of.Inversions") |>
  mutate(AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         Inversions = as.factor(Inversions)) |>
  select(-c(19:21))
coaster
```

#### 4. Take a look at data structure / storage:

```{r}
str(coaster)
```

These look good! All the data types make sense. Taking a look with `summary()`:

```{r}
summary(coaster)
```

#### 5. Document missing values

Seems like there are some NA values in `TopSpeed`, `MaxHeight`, `Drop`, `Length`, and `Duration`. Let's check more thoroughly with `is.na()`:

```{r}
colSums(is.na(coaster))
```

Looks like there are some NAs in the character columns too. Good to know.

### Contingency Tables

#### 6. Basic Tables

Table 1:

```{r}
#| results: hold

cat("=== Table 1: Regions === \n")
table(coaster$Census.Region)

```

This table shows us that the southern census region has the most (223) rollercoasters from this list

```{r}
#| results: hold

cat("=== Table 2: Age Group and Scale ===\n")
table(coaster$AgeGroup, coaster$Scale)
```

This list shows us that there are 147 "Extreme" coasters that are also recently constructed

```{r}
#| results: hold
cat("=== Table 3: Scale, Region, and Inversions ===\n")
table(coaster$Scale, coaster$Census.Region, coaster$Inversions)
```

This three-way table shows us that there are notably fewer coasters with inversions, and that all of them are categorized as "Extreme".

#### 7. Conditional Tables

```{r}
#| results: hold
wooden <- filter(coaster, Type == "Wood")
cat("=== Scale & Design, Wooden Coasters ===\n")
table(wooden$Scale, wooden$Design)
```

This table shows us that there is one singular wooden coaster of design "Bobsled", and that it's scale is "Thrill"

```{r}
#| results: hold
cat("=== Region & Type, Old Coasters ===\n")
oldcoaster <- table(coaster$Census.Region, coaster$Type, coaster$AgeGroup)
oldcoaster[, , 1]
```

#### 8. Tidyverse contingency

```{r}
tidytable <- coaster |> 
  group_by(Census.Region, Scale) |>
  summarize(count = n())

tidytable
tidytable |> pivot_wider(names_from = Census.Region, values_from = count)
```

#### 9. Figures

##### Stacked Bars

```{r}
stacked <- ggplot(data = coaster, aes(x = Census.Region, fill = AgeGroup))+
  geom_bar()+
  labs(x = "Region", y = "Count", title = "Regional coasters by recency of construction")
stacked
```

##### Side-by-side bars

```{r}
sbs <- ggplot(data = coaster, aes(x = Scale, fill = Type))+
  geom_bar(position = "dodge")+
  labs(x = "Excitement Scale", y = "count", title = "Excitement level by material")
sbs
```
