---
title: "HW5_SM"
author: "Sofia Mariano"
format: html
editor: source
---
Prepare session to read-in documents
```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(psych)
```

# Task 1: Querying an API and a Tidy-Style Function
# Task 2: Summarizing Roller Coaster Data
Read-in data set and convert variables to factors as needed
```{r}
coaster <- readr::read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

coaster <- coaster |>
  dplyr::mutate(AgeGroup = as.factor(AgeGroup),
                Scale = as.factor(Scale),
                Type = as.factor(Type),
                Design = as.factor(Design),
                'Inversions?' = as.factor('Inversions?')
                )
```

## Basic EDA
4. Look at data storage and summary.
Data storage:
```{r}
str(coaster)
```

Data summary:
```{r}
psych::describe(coaster)
```

5. Documenting rate of missing values
```{r}
colSums(is.na(coaster))
```

## Categorical Variables
6. Create one-, two-, and three-way contingency tables from one of the variables that was transformed to be a factor variable
one-way contingency table
```{r}
table(coaster$Design)
```
Interpretation: With a frequency of 560/635, this value indicates that sit down roller coasters make up the majority of roller coasters in this dataset.

Two-way contingency table:
```{r}
table(coaster$Design, coaster$Type)
```
Interpretation: A value of 1 indicates that out of the 4 bobsled roller coasters in this dataset, one is made out of wood, making it a less popular building material compared to steel for this type of roller coaster design.

Three-way contingency table:
```{r}
# three-way contingency table
table(coaster$Design, coaster$Type, coaster$AgeGroup)
```
Interpretation: A value of 20 indicates that of the newest inverted roller coasters, all of them were made out of steel compared to wood.

7. Create a conditional two-way table using table() via two different methods.
Method 1: subset the data, then create a two-way table
```{r}
long <- filter(coaster, Duration > 100)
table(long$Design, long$Scale)
```

Method 2: create a three-way table, then subset into a two-way
```{r}
three_way <- table(coaster$State, coaster$AgeGroup, coaster$Scale)
three_way[,,"Extreme"]
```

8. Create a two-way contingency table using dplyr functions.

```{r}
coaster |>
  dplyr::group_by(AgeGroup, Scale) |>
  tidyr::drop_na(AgeGroup, Scale) |>
  dplyr::summarize(count = n())
```

Use pivot_wider to make output look like table() output.
```{r}
coaster |>
  dplyr::group_by(AgeGroup, Scale) |>
  tidyr::drop_na(AgeGroup, Scale) |>
  dplyr::summarize(count = n()) |>
  tidyr::pivot_wider(names_from = AgeGroup, values_from = count)
```

9. Create stacked and side-by-side bar graph.  Give relevant x and y labels and plot titles.
Stacked plot:
```{r}
stacked <- ggplot2::ggplot(data = coaster |>
                           drop_na(AgeGroup, Scale),
                           aes(x = AgeGroup,
                               fill = Scale))
stacked + ggplot2::geom_bar() + 
  ggplot2::labs(x = "Roller Coaster Relative Age Group", 
                y = "Number of Roller Coasters",
                title = "Relative Roller Coaster Age compared to Roller Coaster Scale Rating")
```

Side-by-Side plot:
```{r}
cluster <- ggplot2::ggplot(data = coaster |>
                           drop_na(AgeGroup, Scale),
                           aes(x = AgeGroup,
                               fill = Scale))
cluster + ggplot2::geom_bar(position = "dodge") + 
  ggplot2::labs(x = "Roller Coaster Relative Age Group", 
                y = "Number of Roller Coasters",
                title = "Relative Roller Coaster Age compared to Roller Coaster Scale Rating")
```
