---
title: "HW2_SM"
format: html
editor: visual
---

## Task 1: Basic Vector Practice Create Pre- and Post-procedure Blood Pressure Measurement Vectors.

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Assign names to each vector's elements.

```{r}
patient <- paste("Subject", 1:20, sep = "_")
```

Calculate change in blood pressure following treatment

```{r}
diff_bp <- (pre_bp - post_bp)
```

Average Decrease in BP across all patients

```{r}
mean(diff_bp)
```

Identify patients that experienced a drop in blood pressure following treatment

```{r}
which(diff_bp > 0)
```
Subset patients with blood pressure drop into a vector
```{r}
bp_drop <- diff_bp[diff_bp > 0]

#Confirm subsetting worked by printing vector
bp_drop
```

Calculate average blood pressure drop following treatment
```{r}
mean(bp_drop)
```

## Task 2: Basic Data Frame Practice 
Create initial data frame
```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)

#Confirm dedired data frame created
bp_df
```

Data frame with blood pressure increases post-treatment
```{r}
bp_inc <- bp_df[bp_df$diff_bp < 0, ]

#Confirm desired data frame created
bp_inc
```

Data frame including a column identifying which patients have post_bp < 120
```{r}
bp_low <- (bp_df$post_bp < 120)
bp_all <- data.frame(bp_df,bp_low)

#Confirm desired data frame created
bp_all
```

Print data frame in final document
```{r}
knitr::kable(bp_all)
```

## Task 3: List Practice 
Creation of new data frame to include placebos
```{r}
# Create Vectors of new placebo values
placebo_pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_diff_bp <- (placebo_pre_bp - placebo_post_bp)
placebo_patient <- paste("Subject", 1:10, sep = "_")

# Create Data frame similar to one completed in task 2
placebo_bp_df <- data.frame(placebo_patient, placebo_pre_bp, placebo_post_bp, placebo_diff_bp)
placebo_bp_low <- (placebo_bp_df$placebo_post_bp < 120)
placebo_bp_all <- data.frame(placebo_bp_df, placebo_bp_low)

#Confirm desired data frame created
placebo_bp_all
```

Create list elements corresponding to treatment groups
```{r}
experiment_groups <- list (treatment = bp_all, placebo = placebo_bp_all)

#Confirm desired list created
experiment_groups
```

Accessing first list element with different syntax types
```{r}
experiment_groups[1]
experiment_groups[[1]]
experiment_groups$treatment
```


Access pre_bp column from placebo data frame
```{r}
experiment_groups$placebo$placebo_pre_bp
```

## Task 4: Control Flow Practice
Create new column in both data frames
```{r}
experiment_groups$treatment$status <- character(20)
experiment_groups$placebo$status <- character(10)

#confirm desired columns added to data frames
experiment_groups
```

Create status value for treatment group
```{r}
for (i in 1:20){
  if (experiment_groups$treatment[i,3] <= 120){
    experiment_groups$treatment[i,6] <- 'optimal'
  } else if (experiment_groups$treatment[i,3] <= 130){
    experiment_groups$treatment[i,6] <- 'borderline'
  } else if (experiment_groups$treatment[i,3] > 130){
      experiment_groups$treatment[i,6] <- 'high'
    }
}

#Confirm conditional logic worked
experiment_groups$treatment
```


Create status value for placebo group
```{r}
for (i in 1:10){
  if (experiment_groups$placebo[i,3] <= 120){
    experiment_groups$placebo[i,6] <- 'optimal'
  } else if (experiment_groups$placebo[i,3] <= 130){
    experiment_groups$placebo[i,6] <- 'borderline'
  } else if (experiment_groups$placebo[i,3] > 130){
      experiment_groups$placebo[i,6] <- 'high'
    }
}

#Confirm conditional logic worked
experiment_groups$placebo
```

