---
title: "Homework 3"
author: "Emily Surrell"
format: html
---

### Task 1: Conceptual Questions

1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

    The relative path would be `"../MyData.csv"`. The `../` moves from the working directory `myfolder/homework/` up one level to the directory `myfolder/`.

2.  What are the major benefits of using `R` projects?

    An `R` project is associated with a particular file folder, which becomes the working directory for the project. This works nicely with coding using relative paths. Also, when you render a `.qmd` file, the working directory for that rendering corresponds to the folder in which the `.qmd` file is located.

3.  What is `Git` and what is `GitHub`?

    `Git` is version control software, which is useful for tracking changes to files and collaborating with others. `GitHub` is an online hosting service for `Git`-based projects, allowing your repository to be accessed and stored on the web.

4.  What are the two main differences between a `tibble` and a `data.frame`?

    1.  The default `print()` method for a `tibble` has fancier printing: it shows the number of observations and columns and displays only the first few rows/columns.
    2.  `tibbles` do not coerce down to a vector when you subset to only one column using `[`.

### Task 2: Reading Delimited Data

Read in Glass data (comma-delimited).

```{r}

#bring in tidyverse library
library(tidyverse)

#read in glass data
glass <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type'))

#print glass data
glass
```

Read in Yeast data (space-delimited).

```{r}

#read in yeast data

yeast <- readr::read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
  col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))

#print yeast data
yeast
```

### Task 3: Combining Excel and Delimited Data

Read in first sheet from excel file.

```{r}

#bring in the readxl libarary
library(readxl)

#read in the white wine data
white <- readxl::read_excel("white-wine.xlsx", sheet = 1)

#print the white wine data
white
```

Read in second sheet from excel file.

```{r}

#read in column names for the white wine data
colnamesdata <- readxl::read_excel("white-wine.xlsx", sheet = 2)

#print the column names for the white wine data
colnamesdata
```

Use data from second sheet to overwrite column names from first sheet.

```{r}
#overwrite colnames from white dataset
#t(colnamesdata) transposes colnamesdata from a column to a row
colnames(white) <- t(colnamesdata)

#print data
white
```

Add a column to indicate the wines are white.

```{r}

#assign the value "white" to every observation of a column called type
white$type <- "white"

#print data
white

```

Read in red wine data (semicolon-delimited).

```{r}

#read in red wine data
red <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";", col_types = "ddddddddddddd")

#print data
red
```

Replace the variable names in the red wine data.

```{r}

#overwrite colnames from red dataset
#t(colnamesdata) transposes colnamesdata from a column to a row
colnames(red) <- t(colnamesdata)

#print data
red

```

Append a column denoting the type as red.

```{r}

#assign the value "red" to every observation of a column called type
red$type <- "red"

#print data
red

```

Combine the white and red wine data together.

```{r}
#bring in dplyr library
library(dplyr)

#combine white and red wine data into a single tibble
wine <- dplyr::bind_rows(white, red)

#print data
wine
```
