---
title: "HW3_CM"
author: Cameron Mullaney
date: today
format: html
editor: visual
---

### Task 1: Conceptual Questions

1.  To do this, you need to add "../", so you would need `read_csv("../MyData.csv")`. `../` will take you one step "back" in your working directory, and then you can navigate from there.
2.  R projects function as a storage space for all information associated with a given... project. They have a separate workspace, working directory, and source documents, making it simpler and easier to keep all relevant files in one place.
3.  Git is a version control software that allows multiple people to remotely work on the same project simultaneously. GitHub is the online interface for this software, and it doubles as a public distribution system for those projects.
4.  Tibbles and data frames are very similar, but the main differences are that
    1.  Tibbles do less guessing when used (i.e. no partial matching), so errors come up earlier
    2.  When printing, tibbles limit the number of rows printed to 10

### Task 2: Reading Delimited Data

```{r}
library(tidyverse)
## The glass data has no headers
glassColNames <- c('id', 'ri', 'na', 'mg', 'al', 'si', 'k', 'ca', 'ba', 'fe',
                   'type')
glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                  col_names = glassColNames)

glass
```

```{r}
yeastColNames<-c('seq', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class')

yeast <- read_delim("https://www4.stat.ncsu.edu/online/datasets/yeast.data", 
                    delim = "  ", col_names = yeastColNames)
yeast
```
### Task 3: Combining Excel and Delimited Data
```{r}
## Need the readxl package!
library(readxl)

white <- read_excel("HW_3/white-wine.xlsx", sheet = 1)

```

