---
title: "Homework 3"
author: "ST 558 Group 5"
date: 9/9/2026
format: html
---


### Task 1: Conceptual Questions


1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

    The relative path would be `"../MyData.csv"`. The `../` moves from the working directory `myfolder/homework/` up one level to the directory `myfolder/`.

2.  What are the major benefits of using `R` projects?

    - Allows for the creation of a file hierarchy when working on one project that contains multiple files
    - Associated with a directory, which becomes the working directory for the project
    - Working directory can be used for coding using relative paths
    - When you render a `.qmd` file, the working directory for that rendering corresponds to the folder in which the `.qmd` file is located
    - Connects to `Git` and `Github` to allow for version control and collaboration, which can be managed directly in `RStudio`
    - Functions as a storage space for all information associated with a given project

3.  What is `Git` and what is `GitHub`?

    `Git` is version control software, which is useful for tracking changes to files and collaborating with others. `GitHub` is an online hosting service for `Git`-based projects, allowing your repository to be stored on the web and accessed remotely.

4.  What are the two main differences between a `tibble` and a `data.frame`?

    A `tibble` is a type of `data.frame` that is specific to the `tidyverse` package. The two main differences are:

    1.  The default `print()` method for a `tibble` has fancier printing: it shows the number of observations and columns and displays only the first few rows/columns.
    2.  `tibbles` do not coerce down to a vector when you subset to only one column using `[`.


### Task 2: Reading Delimited Data


Read in Glass data (comma-delimited).

```{r}

#bring in tidyverse library
library(tidyverse)

#read in glass data using read_csv function from the readr package
glass <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type'))

#print glass data
glass
```

Read in Yeast data (space-delimited).

```{r}

#read in yeast data using read_table function from readr package
yeast <- readr::read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
  col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))

#print yeast data
yeast
```


### Task 3: Combining Excel and Delimited Data


Read in first sheet from excel file.

```{r}

#bring in the readxl library
library(readxl)

#read in the white wine data using the read_excel function from the readxl package
white <- readxl::read_excel("white-wine.xlsx", sheet = 1)

#print the white wine data
white
```

Read in second sheet from excel file.

```{r}

#read in column names using the read_excel function from the readxl package
colnamesdata <- readxl::read_excel("white-wine.xlsx", sheet = 2)

#print the column names
colnamesdata
```

Use data from second sheet to overwrite column names from first sheet.

```{r}

#overwrite colnames from white dataset
#t(colnamesdata) transposes colnamesdata from a column to a row
colnames(white) <- t(colnamesdata)

#print data
white
```

Add a column to indicate the wines are white.

```{r}

#assign the value "white" to every observation of a column called type
white$type <- "white"

#see what the new column looks like
head(white$type)

```

Read in red wine data (semicolon-delimited).

```{r}

#read in red wine data using the read_delim function from the readr package
red <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")

#print data
red
```

Replace the variable names in the red wine data.

```{r}

#overwrite colnames from red dataset
#t(colnamesdata) transposes colnamesdata from a column to a row
colnames(red) <- t(colnamesdata)

#print data
red

```

Append a column denoting the type as red.

```{r}

#assign the value "red" to every observation of a column called type
red$type <- "red"

#see what the new column looks like
head(red$type)

```

Combine the white and red wine data together.

```{r}

#bring in dplyr library
library(dplyr)

#combine white and red wine data into a single tibble
wine <- dplyr::bind_rows(white, red)

#print data
wine
```


### Task 4: Writing a Function


Write getIQR function.

```{r}

getIQR <- function(x){
  if(!is.vector(x)){
    stop("Not a vector.")
  }
  return(as.vector(quantile(x)[4] - quantile(x)[2]))
}

```

Use the getIQR function on `fixed_acidity` and `citric_acid`.

```{r}

getIQR(wine$fixed_acidity)
getIQR(wine$citric_acid)

```

Write getCoefVar function.

```{r}

getCoefVar <- function(x){
  if(!is.vector(x)){
    stop("Not a vector.")
  }
  return(sd(x) / abs(mean(x)))
}

```

Use the getCoefVar function on `fixed_acidity` and `citric_acid`.

```{r}

getCoefVar(wine$fixed_acidity)
getCoefVar(wine$citric_acid)

```

Write getSummaries function.

```{r}

getSummaries <- function(x){
  if(!is.vector(x)){
    stop("Not a vector.")
  }
  return(c(IQR = getIQR(x), CoefVar = getCoefVar(x)))
}

```

Use the getSummaries function on `fixed_acidity` and `citric_acid`.

```{r}

getSummaries(wine$fixed_acidity)
getSummaries(wine$citric_acid)

```

Write infix function and test using it.

```{r}
#Write function
`%+%` <- function(a, b){
  return(cbind(a,b))
}

#Create test data
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

#Use the function
df1 %+% df2
df1 %+% df2 %+% vec1

```

