---
title: "Homework 3"
author: Cameron Mullaney
date: today
format: html
---

### Task 1: Conceptual Questions

1.  To do this, you need to add "../", so you would need `read_csv("../MyData.csv")`. `../` will take you one step "back" in your working directory, and then you can navigate from there.

2.  R projects function as a storage space for all information associated with a given... project. They have a separate workspace, working directory, and source documents, making it simpler and easier to keep all relevant files in one place.

3.  Git is a version control software that allows multiple people to remotely work on the same project simultaneously. GitHub is the online interface for this software, and it doubles as a public distribution system for those projects.

4.  Tibbles and data frames are very similar, but the main differences are that

    1.  Tibbles do less guessing when used (i.e. no partial matching), so errors come up earlier
    2.  When printing, tibbles limit the number of rows printed to 10

### Task 2: Reading Delimited Data

First, load in the **tidyverse** package from the library.

Then the `glass` data can be loaded in directly from its URL using `read_csv()` from the `readr` package

```{r}
#| warning: false
#| results: hold

library(tidyverse)
## The glass data has no headers
glassColNames <- c('id', 'ri', 'na', 'mg', 'al', 'si', 'k', 'ca', 'ba', 'fe', 'type')
glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                  col_names = glassColNames)
glass
```

Next, load in the `yeast` data. The data is separated by uneven numbers of spaces, so `read_table()` is used.

```{r}
#| results: hold

yeastColNames<-c('seq', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class')

yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = yeastColNames, show_col_types = FALSE)
yeast
```

### Task 3: Combining Excel and Delimited Data

White wine first - loading data from excel file in repo

```{r}
#| warning: false
## Need the readxl package!
library(readxl)
white <- read_excel("white-wine.xlsx", sheet = 1)
white
```

A couple data cleaning tasks:

```{r}
## Pull in better column names from second sheet of .xlsx file
newNames <- read_excel("white-wine.xlsx", sheet = 2)

## Transpose them, and then overwrite old column names
colnames(white) <- t(newNames)

## Additional column to specify wine color
white$color <- "white"
```

On to red wine!

```{r}
## Read in file from link - *semicolon* separated values
red <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
                delim = ";", show_col_types = FALSE)

## Replace names using the same name list
colnames(red) <- t(newNames)

## Additional column to specify wine color
red$color <- "red"
```

Now, to put it all together! Combining both wine tibbles into one

```{r}
wine <- dplyr::bind_rows(white, red)
wine
```

### Task 4: Writing a Function

Creating `getIQR`, which returns the interquartile range of a provided numeric vector.

```{r}
getIQR <- function(x){
  q <- quantile(x, probs = c(.25, .75))
  iqr <- unname(q[2] - q[1])
  return(iqr)
}
```

Testing out the function

```{r}
cat("Variable \t IQR",
    "\nFixed Acidity:\t", getIQR(wine$fixed_acidity), 
    "\nCitric Acid:\t", getIQR(wine$citric_acid))
```

Seems like it works!

Now, creating a second function, `getCoefVar`, which returns the *sample coefficient of variation*

( sample standard deviation / \| sample mean \| )

```{r}
getCoefVar <- function(x){
  return(sd(x)/abs(mean(x)))
}
```

Testing out the function on the same columns as above

```{r}
cat("Variable \t Coefficient of Variance",
    "\n====================================",
    "\nFixed Acidity:\t", getCoefVar(wine$fixed_acidity), 
    "\nCitric Acid:\t", getCoefVar(wine$citric_acid))
```

Seems like it works!

Now, creating a third function, which utilizes the first two - `getSummaries()` returns a list containing the results of `getIQR()` and `getCoefVar()`

```{r}
getSummaries <- function(x){
  iqr <- round(getIQR(x), 4)
  coefvar <- round(getCoefVar(x), 4)
  
  return(c("IQR" = iqr, "coefVar" = coefvar))
}
```

Using `getSummaries()` to return both calculations in one function

```{r}
#| results: hold

acidity <- getSummaries(wine$fixed_acidity)
cat("=== Fixed acidity ===\n")
acidity

citric <- getSummaries(wine$citric_acid)
cat("\n=== Citric Acid ===\n")
citric
```

"Inventing" a "new" function `%~%` (cbind in disguise)

```{r}
`%~%` <- function(a,b){
  return(cbind(a,b))
}
```

Creating some data to test the "new" function `%~%` on:

```{r}
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
```

```{r}
#| results: hold

test1 <- df1%~%df2
test2 <- df1%~%df2%~%vec1

cat("=== Test 1: ===\n")
test1
cat("\n=== Test 2 ===\n")
test2
```
