---
title: "HW3_SM"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  If your working directory is **myfolder/homework/**, what *relative* path would you specify to get the file located at **myfolder/MyData.csv**?
* "../MyData.csv"

2.  What are the major benefits of using R projects?
* Allows for the creation of a file hierarchy when working on one project that contains multiple files
* It's association with a directory allows for ease in sharing/ collaborating
* R projects can be connected with Git and GitHub to allow for version control and collaboration

3.  What is **Git** and what is **GitHub**?
* Git: a software that allows for version tracking of different projects.
* GitHub: an online hosting service for Git projects that allows for storage and sharing of these projects.

4.  What are the two main differences between a **tibble** and a **data.frame**?
* A tibble is a type of data frame that is specific to the tidy verse package. A data frame cannot exist in the tidyverse if it's not a tibble.
* When a tibble is printed, we get information about the number of observations and see an output of the first few columns. When a data frame is printed, the output is generally the entire data frame regardless of the number of observations.

# Task 2: Reading Delimited Data
Prepare session to read-in data sets
```{r}
library(tidyverse)
```

Read-in Data sets
```{r}
# Reading in glass data
glass.data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = c("ID","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Glass")
                       )

# Print glass data tibble
glass.data

# Reading in yeast data
yeast.data <- read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                       fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                                  c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")
                                  ))

# Print yeast data tibble
yeast.data
```


# Task 3: Combining Excel and Delimited Data

# Task 4: Writing a Function
