---
title: "HW3_SM"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  If your working directory is **myfolder/homework/**, what *relative* path would you specify to get the file located at **myfolder/MyData.csv**?

- "../MyData.csv"

2.  What are the major benefits of using R projects?

- Allows for the creation of a file hierarchy when working on one project that contains multiple files
- It's association with a directory allows for ease in sharing/ collaborating
- R projects can be connected with Git and GitHub to allow for version control and collaboration

3.  What is **Git** and what is **GitHub**?

- Git: a software that allows for version tracking of different projects.
- GitHub: an online hosting service for Git projects that allows for storage and sharing of these projects.

4.  What are the two main differences between a **tibble** and a **data.frame**?

- A tibble is a type of data frame that is specific to the tidy verse package. A data frame cannot exist in the tidyverse if it's not a tibble.
- When a tibble is printed, we get information about the number of observations and see an output of the first few columns. When a data frame is printed, the output is generally the entire data frame regardless of the number of observations.

# Task 2: Reading Delimited Data

Prepare session to read-in data sets

```{r}
library(tidyverse)
```

Read-in Data sets

```{r}
# Reading in glass data
glass.data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = c("ID","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Glass")
                       )

# Print glass data tibble
glass.data

# Reading in yeast data
yeast.data <- read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                       fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                                  c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")
                                  ))

# Print yeast data tibble
yeast.data
```

# Task 3: Combining Excel and Delimited Data

Prepare session to read in and combine excel and delimited data files

```{r}
library(readxl)
library(readr)
library(dplyr)
```

Read in white wine data set

```{r}
#Read in white wine data set
white_wine.data <- read_excel("white-wine.xlsx",
                              sheet = 1
                              )

#Print white wine tibble
white_wine.data

```

Clean white wine data set

```{r}

#Correct variable column names
wine_var_names <- read_excel("white-wine.xlsx",
                              sheet = 2
                           )
colnames(white_wine.data) <- t(wine_var_names)

#Add type column
white_wine.data$type <- "white"

#Print white wine tibble to confirm cleaning and column addition were successful
white_wine.data 
```

Read in red wine data set

```{r}
red_wine.data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                            delim = ";",show_col_types = FALSE)

#Print red wine tibble
red_wine.data
```

Clean red wine data set

```{r}
#Fix Column Names
colnames(red_wine.data) <- t(wine_var_names)

#Add type column
red_wine.data$type <-"red"

#Print red wine tibble to confirm cleaning and column addition were successful
red_wine.data 
```

Combine red and white wine data sets into single tibble

```{r}
wine.data <- dplyr::bind_rows(white_wine.data,red_wine.data)

#Print out wine data to observe if desired results occurred
wine.data
```

# Task 4: Writing a Function

Create IQR function that takes in a vector of numbers and returns interquartile range of that vector

```{r}
getIQR <- function(x){
  p <- quantile(x, probs = c(0.25,0.75))
  IQR <- (p[2]-p[1])
  return(IQR)
}
```

Apply IQR function to fixed_acidity and citric_acid columns from wine.data data set

```{r}
#Apply function to both columns
fixed_IQR <- getIQR(wine.data$fixed_acidity)
citric_IQR <- getIQR(wine.data$citric_acid)

#Print results to see if desired outcome achieved
fixed_IQR
citric_IQR
```

Write a CV function that returns a sample's coefficient of variation.

```{r}
getCoefVar <- function(x){
  return(sd(x)/mean(x))
}
```

Apply CV function to fixed_acidity and citric_acid columns from wine.data data set

```{r}
#Apply function to both columns
fixed_CV <- getCoefVar(wine.data$fixed_acidity)
citric_CV <- getCoefVar(wine.data$citric_acid)

#Print results to see if desired outcome achieved
fixed_CV
citric_CV
```

Create a summary function that takes in a vector and returns both the IQR and the CV of that vector

```{r}
getSummaries <- function(x){
  return(list(IQR = getIQR(x), CV = getCoefVar(x)))
}
```

Apply summary function to fixed_acidity and citric_acid columns from wine.data data set

```{r}
#Apply function to yeast data set
fixed_sum <- getSummaries(wine.data$fixed_acidity)
citric_sum <- getSummaries(wine.data$citric_acid)

# Print results to see if desired outcome achieved
fixed_sum
citric_sum
```

Create own function to replace cbind function

```{r}
#Look at cbind code
cbind

#Create a function that mimics cbind
'%getbind%'<- function(x,y){
  return(cbind(x,y))
}
```

Read-in practice data frames

```{r}
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
```

Check custom function against practice data frames

```{r}
# Check against df1 and df2
practice_1 <- df1%getbind%df2

#Check anticipated results obtained
practice_1

#Check against df1, df2, and vec1
practice_2 <- practice_1%getbind%vec1

#Check anticipated results obtained
practice_2

```
