---
title: "Homework 3"
author: "Emily Surrell"
format: html
---

### Task 1: Conceptual Questions

1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

    The relative path would be `"../MyData.csv"`. The `../` moves from the working directory `myfolder/homework/` up one level to the directory `myfolder/`.

2.  What are the major benefits of using `R` projects?

    An `R` project is associated with a particular file folder, which becomes the working directory for the project. This works nicely with coding using relative paths. Also, when you render a `.qmd` file, the working directory for that rendering corresponds to the folder in which the `.qmd` file is located.

3.  What is `Git` and what is `GitHub`?

    `Git` is version control software, which is useful for tracking changes to files and collaborating with others. `GitHub` is an online hosting service for `Git`-based projects, allowing your repository to be accessed and stored on the web.

4.  What are the two main differences between a `tibble` and a `data.frame`?

    1.  The default `print()` method for a `tibble` has fancier printing: it shows the number of observations and columns and displays only the first few rows/columns.
    2.  `tibbles` do not coerce down to a vector when you subset to only one column using `[`.

### Task 2: Reading Delimited Data

Read in Glass data (comma-delimited).

```{r}

#bring in tidyverse library
library(tidyverse)

#read in glass data
glass <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type'))

#print glass data
glass
```

Read in Yeast data (space-delimited).

```{r}

#read in yeast data

yeast <- readr::read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
  col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))

#print yeast data
yeast
```

